%===============================================================================
\documentclass[man]{apa}%can be jou (for journal), man (manuscript) or doc (document)
\usepackage{url}   %this allows us to cite URLs in the text
\usepackage{graphicx}  %allows for graphic to float when doing jou or doc style
\usepackage{amssymb}  %use formatting tools  for math symbols
%===============================================================================

\usepackage{blindtext}

%===============================================================================
%Here is where we start the important APA stuff
\title{Geo-location, Social Media and Privacy}
\author{Dustin Chilson}
\affiliation{Information Security Intelligence \\ Ferris State University}
\shorttitle{Social Media \& Privacy}
\rightheader{Social Media}
\leftheader{Dustin Chilson}
\note{ISIN-499-Capstone \\ Spring 2011 \\ April 24, 2011}
\abstract{
    The use of social media and location aware applications are on the rise bringing
    with is a host of problems for the privacy of people using these applications.
    This project will look at the uses for location aware applications and social
    media and look at the current or possible uses for the massive amount of
    location data that is collected every day. The project will look at the privacy
    threats that present themselves with the use of these applications and weight
    the utility of these applications against those threats.  
}
%===============================================================================

%===============================================================================
\begin{document} 
\maketitle  
Bla Bla about Intro to Social media 

-Growth as a medium 

-current usage numbers 

-privacy issues 

-data collected

-Explanation of paper

\section{Social Media and Personal Data}
\subsection{Data Collected By Social Media Applications}
General idea of what data is collected and the privacy impact that the
collection may have on the person submitting the data. Explain the specific data
that is collected by site.

\subsubsection{Facebook}
\subsubsection{FourSquare}
\subsubsection{Gowalla}
\subsubsection{Twitter}

\subsection{Data available via APIs}
Explain what an API is and how advertisers can use it to target adds and the
security risks associated with them.\footnote{Test Footnote}

\subsubsection{Facebook}
\subsubsection{FourSquare}
\subsubsection{Gowalla}
\subsubsection{Twitter}

\subsection{Comparison of data submitted vs what is Available via API}
Bla Bla Bla

\newpage
\section{Privacy Policies}}
What is a Privacy policy

legal precedence

comparison of sites

\newpage
\section{Geo-Tagged Photos} 
bla bla bla about how data can be stored in photos, options within applications
and some of the more prevalent uses.

\subsection{Exif Data}
about exif tech

\subsection{Tools to extract}
readers \& Cree.py

\subsection{Location APIs}
desc

\subsubsection{Facebook}
\subsubsection{Flickr}
\subsubsection{Picasa}

\subsection{Recommendations for safety}
Bla

\newpage
\section{Tools for Keeping track of public data}
There are tools that are available for uses to track what data is available about
them as well as limit the data that is collected. As a great security practice
is is very useful to limit the exposure of personal data to the world. Just as
keeping a close eye on ones credit score it is important to keep track of what
data is available on the internet. This section will explain some of the many
tools that are available for users to track the data available on the internet
as well as introduce some tools that can be used to protect their personal data.

\subsection{Tracking Tools}
To keep track of the data that is available on the internet there are several
options available. The tools can be as simplistic as visiting every site that
there is an account associated with a person and looking around and a search
using Google. The tools can also be extremely sophisticated and custom built for
the purpose of tracking social media reputation. 

Reputation is something that many companies and individuals need to be concerned
about. Corporations, elected officials, CEOs, and even private citizens may need
to track the impact of data on the internet. For this reason there are custom
built services that track the data that is mentioned all over the internet and
aggregates it into one place for easy analysis. 

Some applications that will scrape all sites for mentions of include
Radian6\footnote{http://www.radian6.com/products/radian6-dashboard/} and 
Trackur\footnote{http://www.trackur.com/}. These applications search the
scraped version of the internet and applies a filter based on the keywords that
are supplied to the application. 

Radian6 describes itself as a "flexible, web-based social media monitoring and 
engagement platform that lets you view relevant conversations happening around 
your brand and products in real time. We aggregate those conversations – saving 
you lots of legwork – and put them into visuals that make analysis and 
measurement meaningful and actionable."~\cite{Radian6}

This may be far more than what is required for the casual user or job seeker.
There are also open source web applications that aggrigate all of the posts
about, by, and to a user on many of the social media sites. One such project is
called ThinkUp

\subsection{Extensions}
AdBlock
AdBlocking 
Ghostery
etc.

\subsection{Anonimization}
proxies
VPN
Tor

%===============================================================================
\bibliography{paper}
%===============================================================================
\end{document}

